\documentclass[12pt,a4paper]{article}

% ── Packages ──────────────────────────────────────────────────────────────────
\usepackage[margin=2.5cm]{geometry}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{array}
\usepackage{longtable}
\usepackage{hyperref}
\usepackage{fancyhdr}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{tabularx}
\usepackage{multirow}
\usepackage{parskip}

\definecolor{namalgreen}{RGB}{34,85,34}
\definecolor{lightgray}{RGB}{245,245,245}

\hypersetup{
    colorlinks=true,
    linkcolor=namalgreen,
    urlcolor=namalgreen,
    citecolor=namalgreen
}

\titleformat{\section}{\large\bfseries\color{namalgreen}}{}{0em}{}[\titlerule]
\titleformat{\subsection}{\normalsize\bfseries}{}{0em}{}

\pagestyle{fancy}
\fancyhf{}
\lhead{\small CSC-271 -- Database Systems}
\rhead{\small Database System Proposal}
\cfoot{\thepage}
\renewcommand{\headrulewidth}{0.4pt}

% ══════════════════════════════════════════════════════════════════════════════
\begin{document}

% ── Title Page ────────────────────────────────────────────────────────────────
\begin{titlepage}
    \centering
    \vspace*{0.5cm}

    {\LARGE\bfseries Namal University, Mianwali}\\[0.3cm]
    {\large Department of Computer Science}\\[0.8cm]
    \rule{\linewidth}{1pt}\\[0.4cm]
    {\large\bfseries CSC-271 -- Database Systems}\\
    {\large\bfseries Complex Computing Problem -- Milestone 1}\\[0.4cm]
    \rule{\linewidth}{1pt}\\[1.2cm]

    {\Huge\bfseries Database System Proposal}\\[0.5cm]
    {\large\bfseries Project Title: \textit{Namal Criko}}\\[0.3cm]
    {\large Cricket Simulation Trainer -- Database Layer}\\[1.5cm]

    \begin{tabular}{lll}
        \toprule
        \textbf{Name} & \textbf{Roll No.} & \textbf{Email} \\
        \midrule
        M. Jamal Ahmad Khan      & Num-Bscs-2024-51 & \href{mailto:bscs24f51@namal.edu.pk}{bscs24f51@namal.edu.pk} \\
        Qazi M. Auon Farooqi     & Num-Bscs-2024-64 & \href{mailto:bscs24f64@namal.edu.pk}{bscs24f64@namal.edu.pk} \\
        Breera Ijaz              & Num-Bscs-2024-20 & \href{mailto:bscs24f20@namal.edu.pk}{bscs24f20@namal.edu.pk} \\
        \bottomrule
    \end{tabular}\\[1.5cm]

    \textbf{Submission Date:} 12th March, 2026\\[0.3cm]
    \textbf{Instructor:} Asiya Batool\\[0.3cm]
    \textbf{Semester:} Spring 2026\\[0.3cm]
    \textbf{GitHub Repository:} \url{https://github.com/MJamalAhmadKhan/Namal-Criko-Database}
    \vfill
\end{titlepage}

\tableofcontents
\newpage

% ══════════════════════════════════════════════════════════════════════════════
\section{Introduction}

Namal Criko is a web-based cricket simulation trainer developed at Namal University,
Mianwali, as part of CSC-225 (Software Engineering, Fall 2025). It enables players
to learn batting and bowling techniques through interactive browser simulations and
provides administrators with tools to manage users and platform content.

This course project focuses on designing and implementing the \textbf{relational database
layer} that supports all platform features: user management, simulation delivery,
performance tracking, help-desk support, administrative reporting, a Top Players
leaderboard, and a motivational Quote of the Day. A simple local web application
using React and Node.js/Express is also developed to demonstrate the database
visually through a browser interface.

% ══════════════════════════════════════════════════════════════════════════════
\section{Problem Statement}

Namal Criko currently has no database layer. The following gaps must be addressed:

\begin{itemize}[leftmargin=1.5em]
    \item \textbf{No persistent storage} for user accounts, approval statuses,
    or profile data.
    \item \textbf{No simulation catalogue schema} to store techniques, categories,
    assets, and instructional content.
    \item \textbf{No performance records} to track player progress, session durations,
    or learning streaks.
    \item \textbf{No ticket management} to handle help-desk queries and their
    lifecycle states.
    \item \textbf{No reporting store} for administrative analytics such as user
    growth and simulation engagement.
    \item \textbf{No top player data} to display real-world cricket statistics
    as motivational benchmarks for learners.
    \item \textbf{No Quote of the Day} mechanism to deliver rotating motivational
    quotes to players at login.
\end{itemize}

This proposal addresses all gaps through a fully normalised MySQL schema with views,
stored procedures, and a lightweight local web interface for demonstration.

% ══════════════════════════════════════════════════════════════════════════════
\section{Objectives}

\begin{enumerate}[leftmargin=1.5em]
    \item Design a normalised relational schema (minimum 3NF) covering all entities
    within 2nd mile stone  of project .
    \item Implement all tables, constraints, primary keys, foreign keys, and indexes
    in MySQL 8.x.
    \item Enable complete user lifecycle management: registration, admin approval,
    profile updates, suspension, and deletion.
    \item Store and serve the full simulation catalogue (10+ techniques) with
    sub-second read latency on indexed queries.
    \item Log every player--simulation interaction to support performance-tracking
    queries for up to 100+ concurrent users.
    \item Manage the complete help-desk ticket lifecycle with referential integrity
    at all stages.
    \item Store professional cricket player profiles and statistics to power a
    motivational Top Players leaderboard.
    \item Deliver a database-driven Quote of the Day shown to each player on login,
    with a seven-day non-repeat window per player.
    \item Build a local React + Node.js/Express web application connected to MySQL
    via the \texttt{mysql2} driver to demonstrate all database features visually.
    \item Produce a complete ER diagram, data dictionary, and seed dataset.
\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section{Scope}

\subsection{Included in the Database System}

\begin{longtable}{>{\bfseries}p{4cm} p{9.5cm}}
    \toprule
    Module & Database Coverage \\
    \midrule
    \endhead

    User Management &
    Players, administrators, roles, account statuses, and
    email-verification tokens. Supports registration, approval,
    profile management, suspension, and deletion. \\[4pt]

    Simulation Catalogue &
    Categories (Level 1--3 hierarchy), techniques, assets (file paths/URLs),
    step-by-step instructions, key points, and common mistakes. \\[4pt]

    Performance Tracking &
    Access logs (player ID, technique ID, timestamp, duration, completion
    flag), learning streak counters, and progress summaries. \\[4pt]

    Help Desk &
    Tickets (subject, category, status, priority), admin responses,
    and ticket status history. \\[4pt]

    Events \& Announcements &
    Cricket events (title, date, venue, description, archived flag)
    managed by administrators. \\[4pt]

    Top Players &
    Professional player profiles (name, nationality, role, style, debut year)
    linked to a statistics table (matches, runs, average, wickets, economy,
    centuries, fifties). Exposed through ranked views filterable by role
    and format. \\[4pt]

    Quote of the Day &
    Curated quotes table (text, author, category, active flag). A stored
    procedure selects one quote per login, enforces a seven-day non-repeat
    window, and logs impressions per player. \\[4pt]

    Audit \& Security &
    Login attempt logs, admin action logs, and password-reset tokens. \\[4pt]

    Web Demo Layer &
    Local React + Node.js/Express app connected via \texttt{mysql2} to
    demonstrate player dashboard, admin panel, Top Players viewer, and
    Quote of the Day through live API calls. \\

    \bottomrule
\end{longtable}

\subsection{Excluded from Scope}

\begin{itemize}[leftmargin=1.5em]
    \item Real-time multiplayer or AI-opponent features.
    \item Payment processing or external API integrations.
    \item Mobile native application.
    \item Video file storage (assets referenced by URL, not stored as BLOBs).
    \item Live cricket data feeds (Top Players data is manually seeded).
    \item Cloud deployment (localhost only for this milestone).
\end{itemize}

% ══════════════════════════════════════════════════════════════════════════════
\section{Data Collection}

\subsection{Sources of Data}

\begin{enumerate}[leftmargin=1.5em]
    \item \textbf{SRS Document (CSC-225, Dec 2025)} --- primary reference for all
    entities, attributes, and business rules.
    \item \textbf{ICC / PCB coaching resources} --- used to populate the simulation
    technique seed data accurately.
    \item \textbf{Official cricket board statistics} (PCB, ICC) --- source for the
    15+ professional player profiles and career statistics.
    \item \textbf{Curated quotes collection} --- 20+ cricket and training motivational
    quotes from published player interviews and coaching literature.
    \item \textbf{MySQL 8.x reference manual} --- guides data-type, indexing, and
    stored-procedure decisions.
\end{enumerate}

\subsection{Data Collection Methods}

\begin{itemize}[leftmargin=1.5em]
    \item \textbf{Requirements extraction} from SRS functional requirements
    FR-001 through FR-023 to derive entity definitions and cardinality rules.
    \item \textbf{Manual seed data entry} for techniques, player profiles,
    statistics, and quotes for testing and demonstration.
    \item \textbf{Peer schema review} to verify normalisation, constraint
    correctness, and naming consistency.
\end{itemize}

% ══════════════════════════════════════════════════════════════════════════════
\section{Technology Stack}

\begin{longtable}{>{\bfseries}p{3.8cm} p{3.2cm} p{6.3cm}}
    \toprule
    Component & Technology & Justification \\
    \midrule
    \endhead

    DBMS & MySQL 8.x & Industry-standard relational DBMS; free and well-documented. \\[4pt]
    Query Language & SQL & Used for all DDL, DML, views, procedures, and triggers. \\[4pt]
    Back-end & Node.js + Express & Lightweight REST API layer between React and MySQL. \\[4pt]
    & &
    \\[4pt]
    Front-end & React.js & Component-based UI to display database data in a browser. \\[4pt]
    Schema Design & MySQL Workbench & ER diagram creation, DDL forward-engineering, query profiling. \\[4pt]
    Version Control & Git / GitHub & SQL scripts, seed data, and React source all version-controlled. \\[4pt]
    Dev Environment & VS Code + XAMPP & Local MySQL instance; consistent across all team members. \\[4pt]
    Testing & Workbench & SQL query testing and REST endpoint verification. \\

    \bottomrule
\end{longtable}

\subsection{Connecting React to MySQL}

React runs in the browser and \textbf{cannot query MySQL directly}. All requests
pass through the Express API, which executes SQL via \texttt{mysql2} and returns
JSON. The three tiers on localhost are:
\textbf{MySQL (3306)} $\to$ \textbf{Express API (5000)} $\to$ \textbf{React (3000)}.
Install dependencies with \texttt{npm install express mysql2 cors}.

% ══════════════════════════════════════════════════════════════════════════════
\section{System Functionality}

\begin{enumerate}[leftmargin=1.5em, itemsep=4pt]

    \item \textbf{User Registration Storage.}
    Insert player records with full name, unique email, hashed password,
    contact number, and initial status \texttt{PENDING}.

    \item \textbf{Admin Approval Workflow.}
    Update player status to \texttt{ACTIVE} or \texttt{REJECTED}; log
    action in the audit table.

    \item \textbf{Authentication Support.}
    Return hashed password and account status for a given email to support
    bcrypt comparison and approval-gate enforcement in the application layer.

    \item \textbf{Profile Read and Update.}
    Allow updates to mutable fields (name, contact, password hash); keep
    email and creation date immutable at the database level.

    \item \textbf{Simulation Catalogue CRUD.}
    Full create, read, update, and delete on categories (three-level
    hierarchy) and techniques including steps, key points, assets,
    difficulty, and duration.

    \item \textbf{Simulation Access Logging.}
    Insert a log record on every view event: player ID, technique ID,
    start timestamp, duration in seconds, completion flag.

    \item \textbf{Performance Metrics.}
    Stored procedures to return: total simulations accessed, total time
    spent, current learning streak, and completed techniques per player.

    \item \textbf{Help Desk Ticket Lifecycle.}
   Create, read, update, and soft-delete on tickets with status history table; referential integrity
    between ticket, player, and admin records enforced at all stages.

    \item \textbf{Event Management.}
    Create, read, update, and soft-delete cricket events with an archived
    flag; ordered chronologically for player-facing queries.

    \item \textbf{Top Players Leader-board.}
    Store professional player profiles and format-specific statistics.
    Provide  different views of player profiles.

    \item \textbf{Quote of the Day.}
    Table of  quotes with category (\texttt{BATTING},
    \texttt{BOWLING}, \texttt{GENERAL}, \texttt{MINDSET}) and active flags

    \item \textbf{Reporting Views.}
    \begin{itemize}[leftmargin=1.2em]
        \item  Registration counts.
        \item  Techniques by access count.
        \item Average time per category.
        \item  Count and age of open tickets.
        \item Batting ,\ bowling time per player.
    \end{itemize}

    \item \textbf{Data Integrity.}
    Foreign keys, unique constraints on email, NOT NULL on mandatory fields,
    and ENUM types for status fields enforced at the engine level.

    \item \textbf{Bulk Status Operations.}
    Stored procedure to batch-approve or batch-suspend a list of player IDs.

    \item \textbf{Soft Delete.}
    \texttt{deleted\_at} timestamp on user and simulation records for
    logical deletion with historical data retained.

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section{User Classes}

\begin{longtable}{>{\bfseries}p{3cm} p{5.5cm} p{5cm}}
    \toprule
    User Class & Role Description & Database Access Level \\
    \midrule
    \endhead

    Player &
    Learns cricket techniques, tracks own progress, submits tickets,
    views Top Players, and receives Quote of the Day on login. &
    Read: simulation catalogue, top player profiles, own profile, daily
    quote. Write: own performance logs, ticket submissions. \\[6pt]

    Administrator &
    Approves accounts, manages content, resolves tickets, posts events,
    maintains player and quote datasets. &
    Full read/write on all tables except raw password hashes. Executes
    all reporting views and stored procedures. All actions audit-logged. \\

    \bottomrule
\end{longtable}

% ══════════════════════════════════════════════════════════════════════════════
\section{Task Division}

All three members contribute equally across every phase. The table below assigns
one \textbf{primary owner} per deliverable for accountability, while the remaining
two members are co-contributors who implement, test, and review that deliverable
alongside the owner. No member is a passive reviewer --- each person writes code
or SQL in every phase.

\begin{longtable}{p{3.8cm} p{3.5cm} p{3.5cm} p{3.5cm}}
    \toprule
    \textbf{Deliverable} &
    \textbf{M. Jamal} \newline \textit{(Bscs-2024-51)} &
    \textbf{Qazi Auon} \newline \textit{(Bscs-2024-64)} &
    \textbf{Breera Ijaz} \newline \textit{(Bscs-2024-20)} \\
    \midrule
    \endhead

    \textbf{CCP Part 1} \newline
    Project Proposal \&
    GitHub Setup &
    Introduction, Objectives,
    Scope sections \newline
    \textit{(Primary)} &
    Tech Stack, System
    Functionality sections \newline
    \textit{(Primary)} &
    Data Collection, User
    Classes, Conclusion sections \newline
    \textit{(Primary)} \\[8pt]

    \textbf{Part 2} \newline
    Conceptual Model ---
    ER Diagram&
    User Management \&
    Help Desk entities,
    relationships \newline
    \textit{(Primary)} &
    Simulation Catalogue \&
    Performance Tracking
    entities \newline
    \textit{(Co-contributor)} &
    Top Players, Quote of Day,
    Events \& Audit entities \newline
    \textit{(Co-contributor)} \\[8pt]

    \textbf{Part 3a} \newline
    Relational Schema \&
    DDL Scripts&
    DDL for User, Roles,
    Audit, and Ticket tables;
    all FK constraints \newline
    \textit{(Co-contributor)} &
    DDL for Simulation,
    Category, Performance
    Log tables; indexes \newline
    \textit{(Primary)} &
    DDL for Top Players,
    Stats, Quotes, Events,
    Impressions tables \newline
    \textit{(Co-contributor)} \\[8pt]

    \textbf{Part 3b} \newline
    Views, Stored Procedures
    \& Seed Data&
    Procedures:
    \texttt{sp\_get\_top\_players},
    \texttt{sp\_bulk\_status};
    seed: player profiles \newline
    \textit{(Co-contributor)} &
    Procedures:
    \texttt{sp\_get\_performance},
    \texttt{sp\_get\_streak};
    seed: simulations \newline
    \textit{(Co-contributor)} &
    Procedure:
    \texttt{sp\_get\_quote\_of\_day};
    all six reporting views;
    seed: quotes \newline
    \textit{(Primary)} \\[8pt]

    \textbf{Part 4} \newline
    GUI Web Application
    (React + Express)&
    Express REST API:
    all routes for users,
    tickets, auth endpoints \newline
    \textit{(Primary)} &
    React pages: Admin
    Panel, User Management,
    Reporting dashboard \newline
    \textit{(Co-contributor)} &
    React pages: Player
    Dashboard, Top Players,
    Quote of Day display \newline
    \textit{(Co-contributor)} \\[8pt]

    &
    &
    &
    \\

    \bottomrule
\end{longtable}

\begin{longtable}{>{\bfseries}p{4cm} p{9.5cm}}
    \toprule
    Member & Summary of Contributions \\
    \midrule
    \endhead
    M. Jamal Ahmad Khan &
    Proposal (Introduction, Objectives, Scope) $\cdot$ ER entities for User \&
    Help Desk $\cdot$ DDL for User/Ticket/Audit tables $\cdot$ Stored procedures
    for Top Players \& bulk operations $\cdot$ Express REST API layer $\cdot$
    Data dictionary for User/Ticket tables. \\[5pt]
    Qazi M. Auon Farooqi &
    Proposal (Tech Stack, System Functionality) $\cdot$ ER entities for
    Simulation \& Performance $\cdot$ DDL for Simulation/Performance tables
    with indexes $\cdot$ Performance stored procedures $\cdot$ React Admin Panel
    \& reporting dashboard $\cdot$ Data dictionary \& README. \\[5pt]
    Breera Ijaz &
    Proposal (Data Collection, User Classes, Conclusion) $\cdot$ ER entities
    for Top Players, Quotes \& Events $\cdot$ DDL for Top Players/Quotes tables
    $\cdot$ All six reporting views \& Quote of Day procedure $\cdot$ React
    Player Dashboard \& Quote display $\cdot$ Data dictionary for
    views/procedures. \\
    \bottomrule
\end{longtable}

\noindent\textbf{Meeting Cadence:} Weekly sync (Zoom or on-campus) to review
progress, resolve blockers, and update the GitHub Projects task board.

% ══════════════════════════════════════════════════════════════════════════════
\section{Conclusion}

This proposal presents the complete database development plan for Namal Criko.
The system will be built on a normalised MySQL 8.x schema covering user management,
simulation catalogue, performance tracking, help-desk, events, Top Players, and
Quote of the Day. Pre-built views and stored procedures will serve all application
queries, and a lightweight local React web application will demonstrate every
database feature through a live browser interface. The deliverables --- schema,
ER diagram, data dictionary, seed data, and web demo --- form a solid foundation
for the subsequent development phases of the platform.

% ══════════════════════════════════════════════════════════════════════════════
\section{References}

\begin{enumerate}[leftmargin=1.5em]
    \item IEEE Std 830-1998, \textit{IEEE Recommended Practice for Software
    Requirements Specifications}. IEEE, 1998.
    \item Namal Criko SRS, CSC-225 -- Software Engineering, Namal University,
    Mianwali, December 2025.
    \item Namal Criko Project Proposal, CSC-225 -- Software Engineering,
    \item MySQL Workbench Documentation. [Online]. Available:
    \url{https://dev.mysql.com/doc/workbench/en/}
    \item npm, \textit{mysql2 Package}. [Online]. Available:
    \url{https://www.npmjs.com/package/mysql2}
    \item GitHub Repository:
    \url{https://github.com/MJamalAhmadKhan/Namal-Criko-Database}
\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\appendix
\section*{Appendix 1: AI Prompts and Tools Used}
\addcontentsline{toc}{section}{Appendix 1: AI Prompts and Tools Used}

\begin{longtable}{|p{3cm}|p{8cm}|p{3cm}|}
    \hline
    \textbf{AI Tool} & \textbf{Prompt / Purpose} & \textbf{Access Date} \\
    \hline
    \endfirsthead
    \hline
    \textbf{AI Tool} & \textbf{Prompt / Purpose} & \textbf{Access Date} \\
    \hline
    \endhead

    Claude (Anthropic) &
    \textbf{Prompt 1:} ``Generate a complete LaTeX proposal for a database systems
    project based on the Namal Criko cricket simulation trainer.''

    \textbf{Prompt 2:} ``Suggest Task Division to show equal contribution across
    all CCP phases aligned with the course outline.''

    \textbf{Purpose:} Initial LaTeX generation and iterative content refinement. &
    March 2026 \\
    \hline

    Google Gemini &
    \textbf{Prompt3:} ``Proofread and improve clarity of the Problem Statement and
    Task Division sections for academic tone and conciseness.''

    \textbf{Purpose:} Grammar checking and section refinement. & March 2026 \\
    \hline

    GitHub Copilot &
    \textbf{Purpose:} Auto-completing repetitive LaTeX table structures and
    formatting commands. & March 2026 \\
    \hline

\end{longtable}

\subsection*{Declaration}
We hereby declare that all AI tools used have been documented as per course
guidelines. The final content has been reviewed and approved by all team members,
who take full responsibility for its accuracy and originality.

\end{document}
